---
title: "物种的occurance数据"
output: html_notebook
---

### 有"种名"的海藻分布数据

```{r}
suppressPackageStartupMessages(library(tidyverse))
library(obistools)
library(spocc)
library(scrubr)
library(worms)
library(progress)
library(furrr)
```

```{r}
# 载入种-名单
seaweed_species_list_path <- '../../Results/Farming_seaweed_name_list.rds'
seaweed_species_list <- read_rds(seaweed_species_list_path)
```

#### convert FAO names to worms

```{r}
correct_worms <- wormsbynames(seaweed_species_list) 
correct_worms <- correct_worms %>% 
    filter(
        !is.na(valid_name),
        rank == 'Species')

valid_sci_names <- correct_worms$valid_name
```

#### download data


```{r}

dwonloading_batch <- valid_sci_names
pb <- progress_bar$new(
    format = "  downloading :what [:bar] :percent eta: :eta",
    total = length(dwonloading_batch),
    clear = F,
    width = 100
    )
get_occ_data <- function(query, limit = 10){
    
    pb$tick(token = list(what = query))
    occ_data_raw <- occ(query = query, 
                    from = c('gbif', 'obis'),
                    has_coords = T,
                    limit = limit,
                    throw_warnings = F)
    
    return(occ_data_raw)
    
}

occ_data_raw <- map(dwonloading_batch, get_occ_data)

readr::write_rds(occ_data_raw, 
                 '/Volumes/MarSSD/species_occ_download/occ_query.rds')

```

### data cleaning

#### scrubr for data clean

```{r}
# res <- read_rds('/Volumes/MarSSD/species_occ_download/all_species.rds')

clean_species <- function(batch){
    
    batch_df <- occ2df(batch) # convert to tibble
    dis_name <- batch$gbif$meta$opts$scientificName # get query names
    batch_df <- batch_df %>% 
        mutate(dis_name = dis_name) %>% # add auery name
        coord_impossible(lat = 'latitude', 'longitude') %>% # drop impossible spatial data
        coord_imprecise() %>%  # drop points with imprecise spatial coordinate
        coord_incomplete() %>% # drop points of incomplete coordinate
        dplyr::rename(decimalLongitude = longitude,
                      decimalLatitude = latitude) %>% # change column names for 'check_on_land' function
        rowid_to_column() # add id column for 'check_on_land' function
    
    # drop points on land, i.e., museum specimens

    on_land_points <- check_onland(batch_df,
                                   report = T,
                                   offline = F)
    
    if (length(on_land_points) != 0) {
        batch_in_marine <- batch_df %>% 
            filter(!(rowid %in% on_land_points$row)) %>% 
            rename(longitude = decimalLongitude,
                   latitude = decimalLatitude)
    }else {
        batch_in_marine <- batch_df %>% 
            rename(longitude = decimalLongitude,
                   latitude = decimalLatitude)
    }
    
    return(batch_in_marine)
}

# speed up processing using future_map
plan('multisession', workers = 5)
with_progress({
    p <- progressor(steps = length(res))
    
    res_dfs <- future_map(res, ~{
        p()
        clean_species(.x)
        })
})

# save half-way
# write_rds(res_dfs, '/Volumes/MarSSD/species_occ_download/clean_records.rds')
```

#### drop duplicate points

```{r}
# origin_dfs <- read_rds('/Volumes/MarSSD/species_occ_download/clean_records.rds')

drop_dup <- function(df){
    
    recs.dups <- duplicated(df %>% dplyr::select(longitude, latitude))
    occurrences <- df[!recs.dups, ]
    
    return(occurrences)
    
}

clean_dfs <- map(clean_dfs, drop_dup)


clean_dfs <- bind_rows(clean_dfs)
```

#### data check

```{r}
# summarise records for species

clean_dfs %>% 
    split(.$dis_name) %>% 
    map_dbl(.f = ~dim(.x)[1]) %>% 
    stack() %>% 
    pull(values) %>% sum()
```

| values | ind                       |
|-------:|:--------------------------|
|   2262 | Alaria esculenta          |
|    138 | Capsosiphon fulvescens    |
|   2711 | Caulerpa racemosa         |
|     43 | Chondracanthus chamissoi  |
|    776 | Codium fragile            |
|     14 | **Dunaliella salina**     |
|    147 | Eucheuma denticulatum     |
|     91 | Gelidium amansii          |
|   1212 | Gracilaria gracilis       |
|    434 | Gracilariopsis longissima |
|     53 | Kappaphycus alvarezii     |
|   5396 | Laminaria digitata        |
|   2630 | Macrocystis pyrifera      |
|     64 | Monostroma nitidum        |
|   5750 | Palmaria palmata          |
|     22 | **Pyropia columbina**     |
|      7 | **Pyropia tenera**        |
|   2001 | Saccharina japonica       |
|  11882 | Saccharina latissima      |
|    437 | Sargassum fusiforme       |
|   1234 | Ulva prolifera            |
|   1928 | Undaria pinnatifida       |

### save data

```{r}
save_csv <- function(df){
    
    file_name <- unique(df$dis_name)
    
    file_path <- paste0('/Volumes/MarSSD/species_occ_download/', file_name, '.csv')
    write_csv(df, file = file_path)
    
    
}

map(clean_dfs, save_csv)
```
